---
title: "Data607:project1"
author: "Daniel Sullivan"
date: "2/23/2021"
output: html_document
---
get all packages
```{r}
library(tidyverse)
library(openintro)
library(readr)
library(dplyr)
```

## upload the data and put it into a data frame
```{r}
tourney_data <- read_file("tournamentinfo.txt")
chess_table <- data.frame(x=str_extract_all(tourney_data, ".\\d.[|] [A-Z].{0,80}\\r\\n.{0,29}"))
colnames(chess_table)<-"x"
chess_table<-chess_table%>%
  separate(x,c("player number", "name","points", "game1", "game2", "game3", "game4","game5","game6", "game7","state","R"), sep="\\|")
```

###start organizing data by first seperating the round data into numeric and charecter values for opponent number and win/loss/draw/bye values respectivly.
```{r}
chess_table<-chess_table%>%
  cbind(str_split_fixed(chess_table$game1, " ", 2))%>%
  cbind(str_split_fixed(chess_table$game2, " ", 2))%>%
  cbind(str_split_fixed(chess_table$game3, " ", 2))%>%
  cbind(str_split_fixed(chess_table$game4, " ", 2))%>%
  cbind(str_split_fixed(chess_table$game5, " ", 2))%>%
  cbind(str_split_fixed(chess_table$game6, " ", 2))%>%
  cbind(str_split_fixed(chess_table$game7, " ", 2))

colnames(chess_table)<-c("player_number", "name","points", "game1", "game2", "game3", "game4","game5","game6", "game7","state","R", "game1_result", "game1_oponent", "game2_result", "game2_oponent", "game3_result", "game3_oponent", "game4_result", "game4_oponent", "game5_result", "game5_oponent", "game6_result", "game6_oponent", "game7_result", "game7_oponent")

chess_table<-chess_table[c("player_number", "name","points","state","R", "game1_result", "game1_oponent", "game2_result", "game2_oponent", "game3_result", "game3_oponent", "game4_result", "game4_oponent", "game5_result", "game5_oponent", "game6_result", "game6_oponent", "game7_result", "game7_oponent")]

chess_table$R <- str_extract(chess_table$R,"R: .\\d{2,5}")%>%
  str_extract("\\d{3,4}")

chess_table$R<-as.numeric(chess_table$R)
chess_table$player_number<-as.numeric(chess_table$player_number)
chess_table$game1_oponent<-as.numeric(chess_table$game1_oponent)
chess_table$game2_oponent<-as.numeric(chess_table$game2_oponent)
chess_table$game3_oponent<-as.numeric(chess_table$game3_oponent)
chess_table$game4_oponent<-as.numeric(chess_table$game4_oponent)
chess_table$game5_oponent<-as.numeric(chess_table$game5_oponent)
chess_table$game6_oponent<-as.numeric(chess_table$game6_oponent)
chess_table$game7_oponent<-as.numeric(chess_table$game7_oponent)
```


###calculate the average oponent by creating a function and using lapply to apply it across the data frame and insert it at the end.
```{r}
R_func<-function(i){
  g1<-chess_table%>%
    filter(player_number == chess_table[i,7])
  g2<-chess_table%>%
    filter(player_number == chess_table[i,9])
  g3<-chess_table%>%
    filter(player_number == chess_table[i,11])
  g4<-chess_table%>%
    filter(player_number == chess_table[i,13])
  g5<-chess_table%>%
    filter(player_number == chess_table[i,15])
  g6<-chess_table%>%
    filter(player_number == chess_table[i,17])
  g7<-chess_table%>%
    filter(player_number == chess_table[i,19])
  z<-mean(c(g1$R,g2$R,g3$R,g4$R,g5$R,g6$R,g7$R), na.rm = TRUE)
  return(z)
}
```

###trim data so that only values needed are present in CSV file
```{r}

chess_table$avg_op_R<-sapply(chess_table$player_number, R_func)

chess_for_csv<-chess_table[c("player_number", "name","points","state","R", "avg_op_R")]

write.csv(chess_for_csv, "project_1.csv")

```



